---
title: "stringr tutorial"
author: "Alicia Halhed"
date: "13/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an RMarkdown file. Code goes between in the grey blocks, whereas the white area is simply a text editor. 

We are going to load in the tidyverse suite of packages, which includes the `stringr` package. We will nonetheless attach the whole `tidyverse` to have access to the pipe operator `%>%`.
```{r}
library(tidyverse)
```

We will be working with `stringr-data` for this tutorial. There are three sets of strings: `sentences`, `fruit`, and `words`
```{r}
head(sentences)
head(words)
head(fruit)
```

What if you would like to change the case of your letter strings?
```{r}
# change to upper case
fruit %>% str_to_upper
# change to lower case
sentences %>% str_to_lower
# change to sentence sentence format
# first letter of first word capitalized
fruit %>% str_to_sentence
```

Can we put multiple strings into a single string?
```{r}
str_c(fruit, sentences, sep = ": ")
```

How many times does the word "the" appear in our sentences?
```{r}
# in each sentence
sentences %>% 
  str_to_lower %>% # we don't care about the capitalization
  str_count(pattern = "the")
# where in the sentence is "the"?
sentences %>% 
  str_to_lower %>%
  str_locate_all("the") %>% # str_locate identifies first occurrence
  head
# how many start with "The"?
sentences %>%
  str_starts("The") %>%
  sum
```

Where is the letter v in fruit names?
```{r}
fruit %>% str_which("v")
```

How long are the strings?
```{r}
# mean sentence length (characters)
sentences %>%
  str_length %>%
  mean
```

Do any of the fruit contain the word "fruit" in their name?
```{r}
# are there any? yes/no
fruit %>% str_detect("fruit")
# which ones?
fruit %>% str_subset("fruit")
```

The function I typically use the most often is the `word`. This function allows you to extract different "words" from your string, with whatever separator you specify.
```{r}
# let's take the second word of every sentence
sentences %>%
  word(2) # this defaults to separating by spaces
# but we could tell it to separate by a different character
sentences %>%
  word(2, sep = "h")
```

If you are looking for tips on using regular expressions with the stringr package, I would suggest looking at the cheat sheet for stringr (available online).